---
output: 
  github_document:
    toc: TRUE
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

<a href="https://travis-ci.org/dynverse/gillespie"><img src="https://travis-ci.org/dynverse/gillespie.svg" align="left"></a> <a href="https://codecov.io/gh/dynverse/gillespie"> [![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/dynverse/gillespie?branch=master&svg=true)](https://ci.appveyor.com/project/dynverse/gillespie)

# `gillespie`: Gillespie's Stochastic Simulation Algorithm for impatient people.

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%",
  fig.path = "man/figures/",
  message = FALSE,
  dpi = 300
)
set.seed(1)
submission_to_cran <- FALSE
library(tidyverse)
```

**gillespie** is a fast, scalable, and versatile framework for simulating large systems with 
Gillespie's Stochastic Simulation Algorithm (SSA). It is conceptually based on 
[GillespieSSA](https://cran.r-project.org/web/packages/GillespieSSA/index.html),
but rewritten entirely in Rcpp with large scale systems in mind to make
it blazingly fast. The SSA methods currently implemented are: Direct, Explicit
tau-leaping (ETL), and the Binomial tau-leaping (BTL).

## Install

You can install the development version of gillespie from GitHub with

```R
devtools::install_github("dynverse/gillespie", build_vignettes = TRUE)
```

## Examples

The following example models are available:

```{r vignettes, results='asis', echo=FALSE}
walk(
  list.files("vignettes", pattern = "*.Rmd"),
  function(file) {
    title <- 
      read_lines(paste0("vignettes/", file)) %>% 
      keep(~grepl("^title: ", .)) %>% 
      gsub("title: \"(.*)\"", "\\1", .)
    vignette_name <- gsub("\\.Rmd", "", file)
    markdown_name <- gsub("\\.Rmd", ".md", file)
    cat(
      "* ",
      ifelse(submission_to_cran, "", "["),
      title, 
      ifelse(submission_to_cran, "", paste0("](vignettes/", markdown_name, ")")),
      ": ",
      "`vignette(\"", vignette_name, "\", package=\"gillespie\")`\n",
      sep = ""
    )
  }
)
```

```{r rerun_vignettes, include = FALSE}
rerun_vignettes <- FALSE
if (rerun_vignettes) {
  for (file in list.files("vignettes", pattern = "*.Rmd", full.names = TRUE)) {
    cat("Running '", file, "'\n", sep = "")
    rmarkdown::render(file, output_format = "github_document")
  }
}
```

## Latest changes

```{r include = FALSE}
show_link <- ifelse(submission_to_cran, "", "or [NEWS.md](inst/NEWS.md) ")
```
Check out `news(package = "gillespie")` `r show_link` for a full list of changes.

<!-- This section gets automatically generated from inst/NEWS.md, and also generates inst/NEWS -->

```{r news, echo=FALSE, results="asis"}
dynutils::update_news()
cat(dynutils::recent_news())
```
