---
output: 
  github_document:
    toc: TRUE
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

[![Build Status](https://travis-ci.org/rcannood/GillespieSSA2.svg?branch=master)](https://travis-ci.org/rcannood/GillespieSSA2)
[![Build status](https://ci.appveyor.com/api/projects/status/5kw9ynlc1mmehrcj?svg=true)](https://ci.appveyor.com/project/rcannood/gillespiessa2)

# `GillespieSSA2`: Gillespie's Stochastic Simulation Algorithm for impatient people.

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%",
  fig.path = "man/figures/",
  message = FALSE,
  dpi = 300
)
set.seed(1)
submission_to_cran <- FALSE
library(tidyverse)
```

**GillespieSSA2** is a fast, scalable, and versatile framework for simulating large systems with 
Gillespie's Stochastic Simulation Algorithm (SSA). This package is the spiritual successor to 
the GillespieSSA package originally written by Mario Pineda-Krch. 

GillespieSSA2 has the following added benefits: 

* The whole algorithm is run in Rcpp which results in major speed improvements (>100x).
  Even your propensity functions (reactions) are being compiled to Rcpp!
* Parameters and variables have been renamed to make them easier to understand.
* Many unit tests try to ensure that the code works as intended.


The SSA methods currently implemented are: Exact (`ssa_exact()`), Explicit tau-leaping (`ssa_etl()`),
and the Binomial tau-leaping (`ssa_btl()`).

## Install

You can install the development version of GillespieSSA2 from GitHub with

```R
devtools::install_github("rcannood/GillespieSSA2", build_vignettes = TRUE)
```

## Examples

The following example models are available:

```{r vignettes, results='asis', echo=FALSE}
walk(
  list.files("vignettes", pattern = "*.Rmd"),
  function(file) {
    title <- 
      read_lines(paste0("vignettes/", file)) %>% 
      keep(~grepl("^title: ", .)) %>% 
      gsub("title: \"(.*)\"", "\\1", .)
    vignette_name <- gsub("\\.Rmd", "", file)
    markdown_name <- gsub("\\.Rmd", ".md", file)
    cat(
      "* ",
      ifelse(submission_to_cran, "", "["),
      title, 
      ifelse(submission_to_cran, "", paste0("](vignettes/", markdown_name, ")")),
      ":  \n",
      "`vignette(\"", vignette_name, "\", package=\"GillespieSSA2\")`\n",
      sep = ""
    )
  }
)
```

```{r rerun_vignettes, include = FALSE}
rerun_vignettes <- FALSE
if (rerun_vignettes) {
  for (file in list.files("vignettes", pattern = "*.Rmd", full.names = TRUE)) {
    cat("Running '", file, "'\n", sep = "")
    rmarkdown::render(file, output_format = "github_document")
  }
}
```

## Latest changes
Check out `news(package = "GillespieSSA2")` or [NEWS.md](NEWS.md) for a full list of changes.

<!-- This section gets automatically generated from NEWS.md -->

```{r news, echo=FALSE, results="asis"}
cat(dynutils::recent_news())
```
