---
title: "Preparing an SSA run with fastgssa."
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<!-- github markdown built using 
rmarkdown::render("vignettes/preparing_a_run.Rmd", output_format = "github_document")
-->

In order to invoke `ssa()`, the stochastic model needs at least four components:

* the initial state vector (`initial_state`),
* the state-change matrix (`nu`),
* the propensity functions (`propensity_funs`), and
* the final time of the simulation (`final_time`).

The initial state vector defines the population sizes in all the states
at $t = 0$. For example, for a system with two species `prey` and `predators`
where both have an initial population size of 1000, the initial state vector
is defined as `r initial_state <- c(x_prey = 1000, x_predators = 1000)`.

The state-change matrix defines the change in the number of individuals in
each state (rows) as caused by one reaction of a given type (columns).
For example, a system with abovementioned species could have three reactions;
one in which the prey population grows, one in which the predator
population grows by feasting on the prey, and one in which the predator
population diminishes. The matrix could then be defined as follows.
```{r}
nu <- matrix(
  c(
    +1, -1, 0,
    0, +1, -1
  ),
  nrow = 2,
  byrow = TRUE,
  dimnames = list(
    c("x_prey", "x_predators"),
    c("p_prey_up", "p_predation", "p_pred_down")
  )
)
```

The propensity functions define the probabilities that a particular
reaction will occur over the next infinitesimal time interval
$\left[t,t+dt \right]$. In the previous example,
the propensity functions and the corresponding constant parameters
can be defined as follows.
```{r}
propensity_funs <- c(
  "p_prey_up = c1 * x_prey",
  "p_predation = c2 * x_prey * x_predators",
  "p_pred_down = c3 * x_predators"
)
params <- c(c1 = 10, c2 = 0.01, c3 = 10)
```

The simulation can be started by calling the `ssa()` function.
```{r}
library(fastgssa)
out <- 
  ssa(
    initial_state = initial_state,
    propensity_funs = propensity_funs,
    nu = nu,
    params = params,
    method = ssa_direct(),
    final_time = 5,
    census_interval = .001,
    verbose = TRUE
  )
```

```{r fig.width=7, fig.height=5}
ssa_plot(out)
```

The propensity functions should be valid `C++` code. If desired,
a buffer can be used to speed up the calculation of the propensity
values, for example as follows.
```R
propensity_funs_with_buffer <- c(
  "calc1 = c1 * x_prey; p_prey_up = calc1 / (calc1 + 1)",
  "calc2 = c2 * x_prey * x_predators; p_predation = calc2 / (calc2 + 1)",
  "calc3 = c3 * x_predators; p_pred_down = calc3 / (calc3 + 1)"
)
```
